Viscous Transport in Eroding Porous Media

### 2019-08-08 docs/transport/erosion_transport.tex
worked on results section for 100 bodies

We calculate the anomalous diffusion rates by analyzing the tracer
trajectories governed by equation~\eqref{eqn:trajectories} for
geometries that are formed by erosion.  Given a set of trajectories, we
define $\lambda_j(t)$ to be the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Then, particle dispersion is characterized as $\sqrt{\sigma_\lambda^2}$.
At early times, the particles have not explored much of the geometry,
and we expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However,
as the particles pass the grains, their trajectories are altered, and we
expect that $\sigma_\lambda^2 \sim t^\alpha$, with $\alpha < 1$,
indicating that the flow is sub-diffusive.  With respect to erosion, as
the bodies erode, we expect that the channelization of the geometry
results in more trajectories undergoing a ballistic motion, and this
would indicate decrease particle dispersion, meaning a larger anomalous
diffusion rate.

@@

To quantify the tortuosity, we compute the relative velocity
(Figure~\ref{fig:Eroding100tort}(a)) and the local tortuosity
(Figure~\ref{fig:Eroding100tort}(b)) of 1000 tracers placed immediately
to the left of the porous region at $x=-1$.  The porosity at this
instance in time is $62.98\%$ and the maximum velocity at $x=-1$ is
$3.90 \times 10^{-4}$. The initial velocity of the tracers is
qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Compared to
Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is much more
discontinuous.  These discontinuities can be explained by examining the
trajectories of tracers in Figure~\ref{fig:Eroding100tort}(c).  Here,
there are many instances of neighboring trajectories that are deflected
apart from one another as they tend to a stagnation point in the flow,
and this results in trajectories with significantly different lengths.
At this instance in time one the tracers travels 25.5\% farther than it
would have if the bodies had been absent, and the average tracer
travelled ???\% farther than it would have if the bodies were absent.
Therefore, at this instance in time, the tortuosity is $1.???$
\todo[inline]{S-H: Fill in the numbers in this paragraph}

@@

{\color{red}
This case is 100 bodies in a Hagen-Poiseuille flow. The background
flow is 
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right]
\end{align}
with floating flow rate $U$ such that pressure drop is fixed from x = -2 to x = 2.
We record the vorticity of 
fluid in colormap during the bodies erosion in Fig.~\ref{fig:Eroding100vort} and 
measure the gap sizes between bodies in Fig.~\ref{fig:Eroding100gap} at four equispaced times.
Since the vorticity is equivalent to the shear stress and the erosion is faster in the regions where
the magnitude of the vorticity is larger 
so the erosion is getting faster and faster 
from the first snapshot to the forth one in Fig.~\ref{fig:Eroding100vort}.
At t = 0.005 the porosity is 55.54\% and the most gap sizes between bodies are less than 0.015 in 
Fig.~\ref{fig:Eroding100gap}(a). Since some of the bodies are vanished in the fluid, the numner of gaps 
decreases and the gap sizes get larger and larger between the remaining bodies. 
In Fig.~\ref{fig:Eroding100gap}(b) and (c) we still can see small gap sizes are in the majority but 
it is not as concentrated as the one we see in Fig.~\ref{fig:Eroding100gap}(a). 
In Fig.~\ref{fig:Eroding100gap}(d) there are only 254 gaps between 82 bodies and 
the distribution of gap sizes spreads over without any major pack.  
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_50}
\caption{100 bodies with 318 gaps at t= 0.005}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_100}
\caption{99 bodies with 312 gaps at t= 0.01}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_150}
\caption{94 bodies with 294 gaps at t= 0.015}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_200}
\caption{82 bodies with 254 gaps at t= 0.02}
\end{subfigure}
\caption{\label{fig:Eroding100gap} The histograms of gaps in the case of erosion of 100 bodies at four time steps.}
\end{figure}

To study the behavior of fluid in the porous media with multiple bodies, 
we release 1000 tracers equispaced in $y$ from the 
cross section x = -1 and record their trajectories and velocities
until they arrive the end cross section x = 1 
and the porosity of this geometric structure is 62.98 \%. 
To understand the global action of fluid qualitatively, 
we only draw 200 historical trajectories from x = -1 to x = 1 
in Fig.~\ref{fig:Eroding100tort}(a). 
When equispaced tracers are released in a 2D tube without any grains, 
the streamlines are straight 
and the intersection points of streamlines and any cross section are equispaced. 
But it is not the same when there are bodies in the fluid. 
In Fig.~\ref{fig:Eroding100tort}(a)
even the tracers are equispaced in $y$ on the cross section $S$, 
the streamlines have been push up and down by the bodies and
some of them are close together.

%To understand the meaning of jump 
%of $\tau$, we use the data we got from local tortuosity $\tau(y)$
%to find the difference of $\tau$ at the neighboring tracers in Fig.~\ref{fig:Eroding100tort}(c) 
%and draw 50 pairs of neighboring trajectories which have the largest difference of $\tau$
%in Fig.~\ref{fig:Eroding100tort_traj}. These pairs of two neighboring trajectories are splitted by bodies
%then rejoin with each other and the splitting-rejoining behavior causes the difference of $\tau$ with respect to the neighboring trajectories.   
%We also can get the idea why the maximum value of $\tau$ is smaller 
%than the one in the case of 20 bodies . 
%The most of the bodies in this case are smaller than those in the case of 20 bodies.
%So when two neighboring trajectories are splitted on the left side by a smaller body they do not
%travel as far as the two splitted by a bigger body do after they rejoin on the right side. 
}

@@

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/tort_eulerian100}
\caption{\label{fig:Eroding100tort_all} The Eulerian and Lagrangian
approach of tortuosity.  The blue dots are Eulerian approach of
tortuosity and the red stars and square are Lagrangian approach of
tortuosity.  The dash line is the fitting line
$\widehat{T}(\phi)=1-p\ln\phi$ as $p=0.2631$ and the root mean square
deviation is $5.2 \times 10^{-3}$.}
\end{figure}

\begin{figure}[H]
\center
\includegraphics*[width =0.45\linewidth]{./figs/100b_first_moment}
\includegraphics*[width =0.45\linewidth]{./figs/100b_second_moment}
\caption{\label{fig:Eroding100anomalous} The analysis of anomalous
diffusion in 100 bodies with respect to five porosities and the
reference line of 100\% porosity.  The first moment (left) is the mean
length of streamlines and the second moment (right) is the 
\todo[inline]{Slope of the top black line is 0.8715}
}
\end{figure}

@@

>>>>

We calculate the anomalous diffusion rates by analyzing the tracer
trajectories governed by equation~\eqref{eqn:trajectories} for
geometries that are formed by erosion.  Given a set of trajectories, we
define $\lambda_j(t)$ to be the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Then, particle dispersion is characterized as $\sqrt{\sigma_\lambda^2}$.
At early times, the particles have not explored much of the geometry,
and we expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However,
as the particles pass the grains, their trajectories are altered, and we
expect that $\sigma_\lambda \sim t^\alpha$, with $\alpha < 1$,
indicating that the flow is sub-diffusive.  With respect to erosion, as
the bodies erode, we expect that the channelization of the geometry
results in more trajectories undergoing a ballistic motion, and this
would indicate decrease particle dispersion, meaning a larger anomalous
diffusion rate.

@@

To quantify the tortuosity, we compute the relative velocity
(Figure~\ref{fig:Eroding100tort}(a)) and the local tortuosity
(Figure~\ref{fig:Eroding100tort}(b)) of 1000 tracers placed immediately
to the left of the porous region at $x=-1$.  The porosity at this
instance in time is $62.98\%$ and the maximum velocity at $x=-1$ is
$3.90 \times 10^{-4}$. The initial velocity of the tracers is
qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Compared to
Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is much more
discontinuous.  These discontinuities can be explained by examining the
trajectories of tracers in Figure~\ref{fig:Eroding100tort}(c).  Here,
there are many instances of neighboring trajectories that are deflected
apart from one another as they tend to a stagnation point in the flow,
and this results in trajectories with significantly different lengths.
At this instance in time one the tracers travels 25.5\% farther than it
would have if the bodies had been absent, and the average tracer
travelled ???\% farther than it would have if the bodies were absent.
Therefore, at this instance in time, the tortuosity is $1.???$.
\todo[inline]{S-H: Fill in the numbers in this paragraph}

@@

The tortuosity at all porosities from the initial condition until all
the bodies have eroded is in Figure~\ref{fig:Eroding100tort_all}.  The
initial tortuosity is around $T=1.2$ and decays towards $T=1$.  The
tortuosities indicated by the blue dots are calculated using the
Eulerian approach~\eqref{eqn:tortuosity2}, while the red marks use the
Lagrangian approach~\eqref{eqn:tortuosity1}.  Again, the two methods
give comparable results indicating that any recirculation is small and
negligible.  As in the previous examples, the tortuosity depends on the
porosity through a power law model with a small root-mean-squared error
of $5.2 \times 10^{-3}$.  For this example, the logarithm tortuosity
model $\hat{T}(\phi) = 1 - p \ln(\phi)$ achieves a similar
root-mean-squared error.  The tortuosity has one interesting feature
near the end of the simulation where it suddenly increases.  This
behavior can be explained by ... \todo[inline]{S-H: Show me what is
going on here as we discussed.}

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/tort_eulerian100}
\caption{\label{fig:Eroding100tort_all} The Eulerian and Lagrangian
approach of tortuosity.  The blue dots are Eulerian approach of
tortuosity and the red stars and square are Lagrangian approach of
tortuosity.  The dash line is the fitting line
$\widehat{T}(\phi)=\phi^{-p}$ as $p=0.2459$ and the root-mean-square
error is $5.5 \times 10^{-3}$.  We note that the model
$\widehat{T}(\phi) = 1-p\ln(\phi)$, with $p=0.2631$ has a comparable
root-mean-square error of $5.2 \times 10^{-3}$.}
\end{figure}

Next, we use the 1000 tracer trajectories to compute the first
(Figure~\ref{fig:Eroding100anomalous}(a)) and second
(Figure~\ref{fig:Eroding100anomalous}(b)) moments of the streamlines at
various porosities.  At high porosities, the first moment is similar to
the first-moment of trajectories in a empty channel (black dashed line).
At smaller porosities, the first moment is always linear, but larger
slopes indicate that the length of the trajectories are longer at lower
porosities \todo[inline]{Don't understand how this is the case}.  The
onset of anomalous diffusion is observed in
Figure~\ref{fig:Eroding100anomalous}(b). At early times, the particle
spreading is comparable to that of an open channel (dashed curve)
corresponding to the ballistic regime.  However, as the trajectories
traverse the geometry, a transition to a sub-diffusion case, especially
at low porosities, is evident.  The dashed-dotted line corresponds has a
slope less than 1 indicating the long time behavior of the trajectories
\begin{align*}
  \sigma_\lambda \sim t^{0.8715}.
\end{align*}

\begin{figure}[H]
\center
\includegraphics*[width =0.45\linewidth]{./figs/100b_first_moment}
\includegraphics*[width =0.45\linewidth]{./figs/100b_second_moment}
\caption{\label{fig:Eroding100anomalous} The first and second moments of
trajectories passing through an eroding geometry at several porosities.
The dashed black line corresponds to a porosity of 100\%, and the
dashed-dotted line is a line of best fit with slope 0.8715.}
\end{figure}

Finally, using network models for flow in porous media~\cite{}, the
anomalous diffusion rate can be linked to channel widths between solid
bodies.  We construct a Delaunay triangulation with the centers of the
eroding bodies corresponding to the vertices of the triangles.  Then, if
the side of a triangle connects two vertices, we call the bodies
neighbors.  Then, for each pair of neighbors, we measure the distance
between the bodies.  Once a Delaunay triangulation is formed, it is used
for all subsequent time steps until a body completely disappears.  At
this point, a new Delaunay triangulation is formed.

We plot the distribution of the gap sizes at four equally spaced points
in time in Figure~\ref{fig:Eroding100gap}.  At the initial configuration, there
are 100 bodies and 318 pairs of neighbors.  The mean distance between
these neighbors is $???  \times 10^{-???}$ and the variance is $???
\times 10^{???}$.  As the bodies erode, the mean and variance of the
distance between the bodies increases.  Finally, at the final time step,
there are only 82 bodies and 254 pairs of neighbors.
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_50}
\caption{100 bodies with 318 gaps at t= 0.005}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_100}
\caption{99 bodies with 312 gaps at t= 0.01}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_150}
\caption{94 bodies with 294 gaps at t= 0.015}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/gap_hist100_200}
\caption{82 bodies with 254 gaps at t= 0.02}
\end{subfigure}
\caption{\label{fig:Eroding100gap} The distribution of the gaps in the
case of erosion of 100 bodies at four time steps.}
\end{figure}

@@

@@

%To study the behavior of fluid in the porous media with multiple bodies, 
%we release 1000 tracers equispaced in $y$ from the 
%cross section x = -1 and record their trajectories and velocities
%until they arrive the end cross section x = 1 
%and the porosity of this geometric structure is 62.98 \%. 
%To understand the global action of fluid qualitatively, 
%we only draw 200 historical trajectories from x = -1 to x = 1 
%in Fig.~\ref{fig:Eroding100tort}(a). 
%When equispaced tracers are released in a 2D tube without any grains, 
%the streamlines are straight 
%and the intersection points of streamlines and any cross section are equispaced. 
%But it is not the same when there are bodies in the fluid. 
%In Fig.~\ref{fig:Eroding100tort}(a)
%even the tracers are equispaced in $y$ on the cross section $S$, 
%the streamlines have been push up and down by the bodies and
%some of them are close together. 
%
%
%%To understand the meaning of jump 
%%of $\tau$, we use the data we got from local tortuosity $\tau(y)$
%%to find the difference of $\tau$ at the neighboring tracers in Fig.~\ref{fig:Eroding100tort}(c) 
%%and draw 50 pairs of neighboring trajectories which have the largest difference of $\tau$
%%in Fig.~\ref{fig:Eroding100tort_traj}. These pairs of two neighboring trajectories are splitted by bodies
%%then rejoin with each other and the splitting-rejoining behavior causes the difference of $\tau$ with respect to the neighboring trajectories.   
%%We also can get the idea why the maximum value of $\tau$ is smaller 
%%than the one in the case of 20 bodies . 
%%The most of the bodies in this case are smaller than those in the case of 20 bodies.
%%So when two neighboring trajectories are splitted on the left side by a smaller body they do not
%%travel as far as the two splitted by a bigger body do after they rejoin on the right side.

%{\color{red}
%This case is 100 bodies in a Hagen-Poiseuille flow. The background
%flow is 
%\begin{align}
%  \UU(\xx)=U \left[
%  \begin{array}{c}
%    1-y^2 \\ 0
%  \end{array}
%  \right]
%\end{align}
%with floating flow rate $U$ such that pressure drop is fixed from x = -2 to x = 2.
%We record the vorticity of 
%fluid in colormap during the bodies erosion in Fig.~\ref{fig:Eroding100vort} and 
%measure the gap sizes between bodies in Fig.~\ref{fig:Eroding100gap} at four equispaced times.
%Since the vorticity is equivalent to the shear stress and the erosion is faster in the regions where
%the magnitude of the vorticity is larger 
%so the erosion is getting faster and faster 
%from the first snapshot to the forth one in Fig.~\ref{fig:Eroding100vort}.
%At t = 0.005 the porosity is 55.54\% and the most gap sizes between bodies are less than 0.015 in 
%Fig.~\ref{fig:Eroding100gap}(a). Since some of the bodies are vanished in the fluid, the numner of gaps 
%decreases and the gap sizes get larger and larger between the remaining bodies. 
%In Fig.~\ref{fig:Eroding100gap}(b) and (c) we still can see small gap sizes are in the majority but 
%it is not as concentrated as the one we see in Fig.~\ref{fig:Eroding100gap}(a). 
%In Fig.~\ref{fig:Eroding100gap}(d) there are only 254 gaps between 82 bodies and 
%the distribution of gap sizes spreads over without any major pack.  
%}
%
